Place of death in NRS COVID-19 deaths: visualisation for 2022 data

### 2022-06-08 covid_nrs_place_of_death_visualisations_2022.Rmd
Added a few more LA-related plots

theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri")
            ))

@@

```{r}
figure_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL

figure_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

# Home deaths by HB over time: 2020-2022, 4-week average

@@

@@

```{r}
figure_proportion_of_deaths_by_place_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80")
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL

@@

### Home deaths only, HB

```{r}
figure_proportion_of_deaths_at_home_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(merged_proportions_of_deaths_by_place %>% select(year, place_of_death, proportion_scotland = proportion_of_total), by = c("year","place_of_death")) %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(proportion_of_total, proportion_scotland), names_to = "measure", values_to = "proportion_of_total") %>%
  mutate(measure = if_else(measure == "proportion_of_total", true = "Health board", false = "Scotland")) %>%
  ggplot(., aes(x=year, y=proportion_of_total, linetype = measure)) +
  geom_point() +
  geom_line() +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80")
    ) +
  labs(
    linetype = NULL,
    x = NULL, y= "Proportion of yearly deaths that happened at home or other non-institution (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL

@@

>>>>

theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri"),
                  axis.text = element_text(size = 9)
            ))

@@

```{r}
(figure_proportion_annual_deaths_by_place <-
  merged_proportions_of_deaths_by_place %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = sort(unique(merged_proportions_of_deaths_by_place$year))) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  # theme_minimal() +
  theme(
    legend.position = "top", 
    plot.caption = element_text(size = 10, colour = "gray60")
    ) +
  labs(
    x = "Year", y = "Proportion of yearly deaths (%)",
    colour = NULL,
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c(death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)
figure_proportion_annual_deaths_by_place %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_annual_deaths_by_place", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 10, dpi = 300, units = "cm")
```

@@

# Home deaths by geography over time: 2020-2022, 4-week average

## Home deaths by HB, 4-week average

@@

## Home deaths by LA, 4-week average

```{r}
(figure_home_deaths_over_time_by_la_4_week_average <-
  merged_deaths_la %>%
  filter(place_of_death == "Home & other non-institution") %>%
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  pivot_longer(cols = c(ma4w_deaths_all_causes, ma4w_deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks", limits = c(earliest_date, most_recent_date_available_2022)) +
  geom_ribbon(aes(ymin = ma4w_deaths_min, ymax = ma4w_deaths_max, fill = "#4477aa"), alpha = 0.5) +
  # geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = ma4w_deaths_mean, linetype = "dashed"), size = 0.2) +
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2021), linetype = "dashed", colour="grey50") +  # start of 2021
  geom_vline(xintercept = compute_start_date_from_week_number(week_number = 1, year_number = 2022), linetype = "dashed", colour="grey50") +  # start of 2022
  # scale_x_continuous(breaks = seq(0,60,by=10)) +
  facet_wrap(~ref_area, scales = "free_y") +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_linetype_manual(values = c("dashed"), labels = "Weekly average") +
  scale_fill_manual(values = c("#4477aa"), labels = "Historic weekly minima/maxima") +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
    ) +
  labs(x = NULL,
       y = "Deaths (average of past 4 weeks)",
       title = "Deaths at Home*: 4-week average by Scottish Local Authority",
       subtitle = date_range_text,
       caption = paste0(c(
         "*Home or other non-institutional setting", vertical_line_text, vertical_scale_differs_between_panels_text,death_registrations_text,source_nrs_text
         ), collapse="\n")
       ) +
  ## annotate historic range
  annotate(geom = "text", x=ymd("2021-01-07"), y=0, label = "2021", hjust = 0, vjust = -1, colour = "grey40") +
  annotate(geom = "text", x=ymd("2022-01-07"), y=0, label = "'22", hjust = 0, vjust = -1, colour = "grey40") +
  NULL
 )

figure_home_deaths_over_time_by_la_4_week_average %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_home_deaths_4week_average_over_time_by_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 40, height = 20, dpi = 300, units = "cm")
```

@@

```{r}
(figure_proportion_of_deaths_by_place_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

## Proportion of deaths by place, by LA

```{r}
(figure_proportion_of_deaths_by_place_by_la <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "la") %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point(aes(size = place_of_death)) +
  geom_line(aes(size = place_of_death)) +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  guides(colour = guide_legend(nrow=2, title = NULL), 
         size = guide_legend(nrow=2, title = NULL)) +
  labs(
    x = NULL, y= "Proportion of yearly deaths (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Local Authorities sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

figure_proportion_of_deaths_by_place_by_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_of_deaths_by_place_by_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 30, height = 25, dpi = 300, units = "cm")
```

## Home deaths only, HB

```{r}
(figure_proportion_of_deaths_at_home_by_hb <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "hb") %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(merged_proportions_of_deaths_by_place %>% select(year, place_of_death, proportion_scotland = proportion_of_total), by = c("year","place_of_death")) %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_hb_by_population)) %>%  # arrange hb/la by population size
  pivot_longer(cols = c(proportion_of_total, proportion_scotland), names_to = "measure", values_to = "proportion_of_total") %>%
  mutate(measure = if_else(measure == "proportion_of_total", true = "Health board", false = "Scotland")) %>%
  ggplot(., aes(x=year, y=proportion_of_total, linetype = measure)) +
  geom_point() +
  geom_line() +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  labs(
    linetype = NULL,
    x = NULL, y= "Proportion of yearly deaths that happened at home or other non-institution (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Health boards sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

@@

## Home deaths only, LA

```{r}
(figure_proportion_of_deaths_at_home_by_la <-
  proportions_of_deaths_by_place_by_geography %>%
  filter(geography == "la") %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(merged_proportions_of_deaths_by_place %>% select(year, place_of_death, proportion_scotland = proportion_of_total), by = c("year","place_of_death")) %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(ref_area = factor(x = ref_area, levels = order_la_by_population)) %>%  # arrange la/la by population size
  pivot_longer(cols = c(proportion_of_total, proportion_scotland), names_to = "measure", values_to = "proportion_of_total") %>%
  mutate(measure = if_else(measure == "proportion_of_total", true = "Health board", false = "Scotland")) %>%
  ggplot(., aes(x=year, y=proportion_of_total, linetype = measure)) +
  geom_point() +
  geom_line() +
  geom_text(data = ~filter(.x, year==2022), aes(x=2022.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  facet_rep_wrap(~ ref_area, repeat.tick.labels = TRUE) +
  scale_size_manual(values = c(1,1.8,1,1), name = "Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  scale_x_continuous(breaks = sort(unique(proportions_of_deaths_by_place_by_geography$year))) +
  scale_y_continuous(label = function(x) scales::percent(x, accuracy = 1)) +
  scale_colour_viridis_d(option="C", name="Place of death", labels = order_of_place_of_death_levels[c(1,3,2,4)]) +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 12),
    plot.caption = element_text(size = 10, colour = "gray60"),
    panel.border = element_rect(fill = NA, colour = "gray80"),  # draw a transparent rectangle as the border
    strip.background = element_rect(fill = NA, colour = "gray80"),
    axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
  labs(
    linetype = NULL,
    x = NULL, y= "Proportion of yearly deaths that happened at home or other non-institution (%)",
    subtitle = paste0("* ",only_2022_text),
    caption = paste0(c("Local Authorities sorted by 2020 population estimate.",death_registrations_text,source_nrs_text),collapse="\n")
    ) +
  NULL)

figure_proportion_of_deaths_at_home_by_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_of_deaths_at_home_by_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 30, height = 25, dpi = 300, units = "cm")
```

## Map-based visualisations

```{r}
## merge the 2015-2019 period historical weekly averages to 2022 data
merged_deaths_la_historic_period_2015_2019 <-
  bind_rows(
    merged_deaths_la %>% filter(year %in% 2020:2021),
    merged_deaths_la %>% filter(year == 2022) %>%
      select(-c(deaths_mean, deaths_min, deaths_max)) %>%
      mutate(historic_period = "2015-2019") %>%
      left_join(historical_weekly_average_la, by = c("week_number", "place_of_death", "ref_area", "historic_period"))
  ) %>%
  left_join(weekly_covid_cases_la, by = c("year","week_number","date_w_c","week_number_run_over","ref_area"))
```

### Home deaths increase since 2019 by LA

For this we can take the total annual deaths in each LA and compare to historical figures, showing the ratio for each year so far; alternatively we can also average over 2020-2022, and compare that to the average!
Because we only have partial data for 2022, we need to employ the weekly comparison method -

```{r}
(figure_map_home_deaths_increase_from_2020_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority", width=40),
       # subtitle = str_wrap("A ratio of 1 means there was no increase; A ratio of 1.6 means there was a 60% increase of home deaths in 2020 over the mean 2015-2019.", width = 80),
       # caption = paste0(c(areas_relatively_higher_number_home_deaths_text, home_deaths_ratio_text, source_nrs_text), collapse="\n"),
       caption = paste0(c(source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

figure_map_home_deaths_increase_from_2020_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_map_home_deaths_increase_from_2020_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 14, dpi = 300, units = "cm")
```

### Home deaths increase since 2019 by LA, by year

For this we can take the total annual deaths in each LA and compare to historical figures, showing the ratio for each year so far; alternatively we can also average over 2020-2022, and compare that to the average!
Because we only have partial data for 2022, we need to employ the weekly comparison method -

```{r}
(figure_map_home_deaths_increase_by_year_la <-
  merged_deaths_la_historic_period_2015_2019 %>%
  filter(historic_period == "2015-2019") %>%
  group_by(year, ref_area, place_of_death) %>%
  summarise(
    ratio_deaths_to_historic = sum(deaths_all_causes, na.rm=TRUE) / sum(deaths_mean, na.rm=TRUE),
    percent_increase = ratio_deaths_to_historic - 1,
    .groups = "drop"
  ) %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase), colour="grey10") +
  geom_sf(colour="grey10") +
  # scale_fill_viridis_b(limits = c(minimum_ratio_between_hb_and_la_home_deaths-1, maximum_ratio_between_hb_and_la_home_deaths-1), labels = function(x) scales::percent(x, accuracy = 1)) +
  scale_fill_viridis_b(limits = c(0, 0.6), labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(
    axis.text = element_blank(), 
    panel.grid = element_blank(),
    # legend.position = "bottom"
    ) +
  facet_wrap(~year, nrow = 1) +
  labs(x="", y="",
       title = "Increased home deaths since the start of 2020 relative to historical mean 2015-2019, by Local Authority",
       caption = paste0(c(source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease relative\nto 2015-2019"
       ) +
  NULL)

figure_map_home_deaths_increase_by_year_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_map_home_deaths_increase_by_year_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 32, height = 18, dpi = 300, units = "cm")
```

### Home deaths increase vs. covid cases by LA, all deaths since 2019

```{r}
# TODO: continue working on this!
(figure_map_covid_cases_from_2020_la <-
  weekly_covid_cases_la %>%
  group_by(ref_area) %>%
  summarise(
    ratio_covid_cases = max(crude_rate_positive_per100k),
    .groups = "drop"
  ) %>% 
  # left_join(population_estimate_la_2019, by = c("ref_area"="la")) %>%
  # dplyr::group_by(ref_area) %>%
  # dplyr::summarise(
  #   ratio_covid_cases = sum(weekly_positive, na.rm=TRUE) / all_ages *1e5, 
  #   .groups = "drop"
  # ) %>%
  left_join(shapefile_la, by=c("ref_area"="LAD20NM")) %>%
  ggplot(aes(geometry=geometry, fill=ratio_covid_cases), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B", labels = function(x) format(x, big.mark=",")) +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  labs(x="", y="",
       title = str_wrap("Total COVID-19 cases since the beginning of 2020 per 100,000 population, by Local Authority", width=50),
       subtitle = "",
       caption = paste0(c(source_phs_text), collapse="\n"),
       fill = str_wrap("Cases per 100,000 population", width = 15)
       ) +
  NULL)

figure_map_covid_cases_from_2020_la %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("figure_map_covid_cases_from_2020_la", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png"), plot = ., width = 13, height = 14, dpi = 300, units = "cm")
```
